R code for 'Apparent physiological costs of blood parasite infection in the early-life of a vertebrate host - only in highest infections

### 2022-02-23 MS_physiology_infection.Rmd
Formatting

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, strip.white = TRUE, collapse = TRUE, fig.align = "center",
                      warning = F, message = F, highlight = TRUE, tidy = TRUE)

# Routine clean-up upon start

rm(list=ls())
```

@@

A number of packages, all of which are available on CRAN are required to repeat analyses presented in this document. The code chunk below allows to load them for the analyses.

```{r packages, message=F, warning=F}
library(chron)
library(cowplot)
library(EnvStats)
library(lme4)
library(lsmeans)
library(ggalt)
library(ggdist)
library(ggeffects)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(ggridges)
library(ggsci)
library(ggsignif)
library(grid)
library(gridExtra)
library(magrittr)
library(patchwork)
library(plyr)
library(rptR)
library(RVAideMemoire)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(tidybayes)
library(tidyverse)
library(wesanderson)
library(car)
theme_set(theme_tidybayes() + panel_border())
```

@@

The raw data is comprised of 276 common buzzard (*Buteo buteo*) individuals. Each individual was sampled twice over the course of nestling period during the years 2016 and 2018 - 2020. First, the raw data is loaded from the file `physiology_infection_stage_full_wideformat.csv` and formatted (scaling variables, reordering factor levels etc.) for downstream analyses

@@

## Arrange data frame
## ----------------------------------------------------------------------------- 
df <- arrange(df, df$Diff_Inf_Class)

```

@@

Investigating treatment effects on the prevalence of *Leucocytozoon* parasites:

```{r, results='hide', message=F, warning=F,prompt=F}

## Build contingency table: Infection prevalence and treatment: First sampling
## -----------------------------------------------------------------------------
pre_treatment <- with(df, table(Prevalence1, Treatment))
prop.test(pre_treatment)
pairwise.prop.test(pre_treatment)

@@

##  Create Figure 2a
## -----------------------------------------------------------------------------
treatment_prevalence_changes  <- ggplot(tab_cont) +  
  aes(Timepoint, Count, fill = Prevalence) +  
  facet_wrap(facets = "Treatment") +  
  geom_bar(stat = 'identity',  width = 0.3) +  
  labs(y = "Count of nestlings") +  
  geom_text(aes(label = Count),position  = position_stack(0.5)) +  
  scale_fill_manual(values = c("#D0D0D0","#A8A8A8")) +  
  theme(axis.title.x  =  element_blank(), 
        strip.text = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16))

@@

## Figure 2b
## -----------------------------------------------------------------------------
treatment_meaninf_changes = ggplot(mean_infint) +
  aes(Sampling_rank, Infection_intensity, group = Treatment) +
  geom_vline(xintercept  =  "First sampling", linetype = "dashed", color = 'darkgrey') +
  geom_vline(xintercept = "Second sampling", linetype = "dashed", color = 'darkgrey') +
  labs(y = "Mean infection intensity") +
  geom_point(aes(colour = Treatment),position = position_jitterdodge(jitter.height  =  0.1))+
  stat_summary(fun = mean, geom = "line",size = 2,color = c("lightgrey","black","black","lightgrey"))+ 
  stat_summary(fun = mean, geom = "point",size = 5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.04, size = 2) + 
  scale_color_manual(values = c("lightgrey","black")) + 
  theme_classic() + 
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16),
        plot.margin = unit(c(0,0,0,30), "pt"))

## Panel wrapping Fig2 a & b:
## ----------------------------------------------------------------------------- 
Fig2 <- treatment_prevalence_changes + 
  treatment_meaninf_changes +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 16, face = "bold"))
ggsave(Fig2,filename = "./Figs/Fig2_efficiency_malarone.png",units = "px", width = 6000, height = 3000)

@@

## Testing for Treatment effect on parasitemia change between first and second sampling
## -----------------------------------------------------------------------------

lmerTest::lmer(intensity_change~Treatment+Year+days_resample+(1|Nest), data=df[!df$intensity_change %in% -100,]) %>%
  summary()
```

@@

```{r}
lm_d1 = lm(Breathing_rate ~ Treatment, data = df)
summary(lm_d1)
```

@@

```{r}
lm_d2 <- lm(body_temperature~Treatment, data = df)
summary(lm_d2)
```

@@

```{r}
lm_d3 <- lm(delta_body_condition~Treatment, data = df)
summary(lm_d3)
```

@@

```{r}
lm_d4 <- lm(Growth_rate~Treatment, data = df)
summary(lm_d4)

@@

Exploring differences in physiological measures that are associated with the infection stage (increase, peak or decreasing) of tested individuals.

@@

## Plotting predicted values across infection stages
## -----------------------------------------------------------------------------
gg1_pred <- ggplot(mod_br) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_br$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Breathing rate") + 
  ylim(c(min(mod_br$conf.low) - 1.5,max(mod_br$conf.high) + 1.5)) + 
  geom_errorbar(aes(ymin = mod_br$conf.low,ymax = mod_br$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br$predicted - mod_br$std.error, ymax = mod_br$predicted + mod_br$std.error),
                size = 8, colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        plot.margin = margin(0, 0, 0, 0, "cm"), 
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
  aes(x = 1, y = max(mod_br$conf.high) + 1, xend = 4, yend = max(mod_br$conf.high) + 1)) + 
  geom_text(aes(x = 2.5, y = max(mod_br$conf.high) + 1.5),
  label = "n.s.", size = 5)
```

@@

# Plotting predicted values across infection stages
gg2_pred <- ggplot(mod_bt) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_bt$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Body temperature") + 
  geom_errorbar(aes(ymin = mod_bt$conf.low,ymax = mod_bt$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt$predicted - mod_bt$std.error,ymax = mod_bt$predicted + mod_bt$std.error),
                size = 8,colour = wes_palette("Zissou1",4), width = 0) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title.y = element_text(size = 14, margin = unit(c(0,3,0,0),"mm")), 
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 40.6, xend = 4, yend = 40.6)) + 
    geom_text(aes(x = 2.5, y = 40.7),
              label = "n.s.", size = 5)
```

@@

gg3_pred = ggplot(mod_bc) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_bc$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Δ Body condition") + 
  geom_errorbar(aes(ymin = mod_bc$conf.low,ymax = mod_bc$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bc$predicted - mod_bc$std.error,ymax = mod_bc$predicted + mod_bc$std.error),
                size = 8,colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bc$conf.high) + 17, xend = 4, yend = max(mod_bc$conf.high) + 17)) + 
  geom_text(aes(x = 2.5, y = max(mod_bc$conf.high) + 23),
              label = "n.s.", size = 5)
```

@@

```{r mixed model growth rate}
# Model 
lmer_gr = lmer(Growth_rate ~ Diff_Inf_Class + Diff_Inf_Class:Treatment + Sex + days_resample + Age_in_days + Year + (1|Nest), data= df)

@@

gg4_pred = ggplot(mod_gr) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_gr$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Growth rate") + 
  geom_errorbar(aes(ymin = mod_gr$conf.low,ymax = mod_gr$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_gr$predicted - mod_gr$std.error,ymax = mod_gr$predicted + mod_gr$std.error), 
                size = 8,colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none", 
        axis.title.y = element_text(size = 14, margin = unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 30, "pt"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_gr$conf.high) + 1.2, xend = 4, yend = max(mod_gr$conf.high) + 1.2)) + 
    geom_text(aes(x = 2.5, y = max(mod_gr$conf.high) + 2),
              label = "n.s.", size = 5)
```

@@

The code below is used to summarise the model output and create Figure 4 and Table 3 as shown in the manuscript.

```{r}
## Figure 4 
## -----------------------------------------------------------------------------
fig4 <- gg1_pred + gg2_pred + gg3_pred + gg4_pred + plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 16, face = "bold"))
ggsave(fig4,filename = "./Figs/Fig4_predicted_physio_param_infection_stages.png",units = "px", width = 3700, height = 3700)

@@

Exploration of effects by parasitemia levels on physiology of individuals.

@@

```{r mixed model breathing rate 2}
## Model 
## -----------------------------------------------------------------------------
lmer_br_para = lmer(
  Breathing_rate ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample + 
    Age_in_days + Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest), data= df)

# Predicted values for post-treament parasitemia

mod_br_para=get_model_data(lmer_br_para,type = 'pred',terms = "Infection_intensity.2") %>% 
   mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

# Table with sample sizes specific for breathing rate change data

m1 = table(df$Infection_intensity.2[!is.na(df$Breathing_rate)])

mod_br_para <- mod_br_para %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",m1)), levels = c(paste0(x,"\nn= ",m1))))
mod_br_para$lower_CI= c(31.11, mod_br_para$predicted[1] - 2.96, mod_br_para$predicted[1] - 6.49,mod_br_para$predicted[1] - 4.67, mod_br_para$predicted[1] - 4.63)
mod_br_para$higher_CI= c(41.42,mod_br_para$predicted[1]+3.90,mod_br_para$predicted[1]+1.75,mod_br_para$predicted[1]+1.46,mod_br_para$predicted[1]+4.45)
mod_br_para$se= c(2.61,1.74,2.09,1.56,2.31)

@@

```{r mixed model growth rate 2}
# Model 
lmer_gr_para = lmer(
  Growth_rate ~ Infection_intensity.2 + Infection_intensity.2:Treatment +
    days_resample + Sex + Age_in_days + Year + (1|Nest), data= df)

# Predicted values for post-treament parasitemia
mod_gr_para=get_model_data(lmer_gr_para,type = 'pred',terms = "Infection_intensity.2") %>% 
   mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

# Table with sample sizes specific for growth rate data

m4 = table(df$Infection_intensity.2[!is.na(df$Growth_rate)])

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 8, fig.height = 8, comment = "#>",
  echo = TRUE, strip.white = TRUE, collapse = TRUE, fig.align = "center",
  warning = F, message = F, highlight = TRUE, tidy = TRUE)
```

@@

A number of packages, all of which are available on CRAN are required to repeat analyses presented in this document. The code chunk below allows to load them for the analyses. Any missing package needs to be installed before, using the 'install.packages())' function:

```{r packages, message=F, warning=F}
library(chron)
library(cowplot)
library(EnvStats)
library(lme4)
library(lmerTest)
library(lsmeans)
library(ggalt)
library(ggdist)
library(ggeffects)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(ggridges)
library(ggsci)
library(ggsignif)
library(grid)
library(gridExtra)
library(magrittr)
library(patchwork)
library(plyr)
library(rptR)
library(RVAideMemoire)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(tidybayes)
library(tidyverse)
library(wesanderson)
library(car)
## set theme for ggplot objects
theme_set(theme_tidybayes() + panel_border())
```

@@

The raw data is comprised of 276 common buzzard (*Buteo buteo*) individuals. Each individual was sampled twice over the course of the nestling period during the years 2016 and 2018 - 2020. First, the raw data is loaded from the file `physiology_infection_stage_full_wideformat.csv` and re-formatted (scaling variables, reordering factor levels etc.) before downstream analyses:

@@

## Arrange data frame
## ----------------------------------------------------------------------------- 
df <- arrange(df, df$Diff_Inf_Class)
```

@@

Investigating treatment (*control* vs *Malarone*) effects on the prevalence of *Leucocytozoon* parasites:

```{r, results='hide', message=F, warning=F,prompt=F}
## Build contingency table: Infection prevalence and treatment: First sampling
## -----------------------------------------------------------------------------
pre_treatment <- with(df, table(Prevalence1, Treatment))
prop.test(pre_treatment)
pairwise.prop.test(pre_treatment)

@@

##  Create Figure 2a
## -----------------------------------------------------------------------------
Fig_2A  <- ggplot(tab_cont) +  
  aes(Timepoint, Count, fill = Prevalence) +  
  facet_wrap(facets = "Treatment") +  
  geom_bar(stat = 'identity',  width = 0.3) +  
  labs(y = "Count of nestlings") +  
  geom_text(aes(label = Count),position  = position_stack(0.5)) +  
  scale_fill_manual(values = c("#D0D0D0","#A8A8A8")) +  
  theme(axis.title.x  =  element_blank(), 
        strip.text = element_text(size = 16, face = "bold"),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16))
Fig_2A

@@

## Figure 2b
## -----------------------------------------------------------------------------
Fig_2B = ggplot(mean_infint) +
  aes(Sampling_rank, Infection_intensity, group = Treatment) +
  geom_vline(xintercept  =  "First sampling", linetype = "dashed", color = 'darkgrey') +
  geom_vline(xintercept = "Second sampling", linetype = "dashed", color = 'darkgrey') +
  labs(y = "Mean infection intensity") +
  geom_point(aes(colour = Treatment),position = position_jitterdodge(jitter.height  =  0.1)) +
  stat_summary(fun = mean, geom = "line",size = 2,color = c("lightgrey","black","black","lightgrey")) + 
  stat_summary(fun = mean, geom = "point",size = 5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.04, size = 2) + 
  scale_color_manual(values = c("lightgrey","black")) + 
  theme_classic() + 
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16),
        axis.title.y = element_text(size = 18, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 16),
        plot.margin = unit(c(0,0,0,30), "pt"))

## Panel wrapping Fig2 a & b:
## ----------------------------------------------------------------------------- 
Fig2 <- Fig_2A + Fig_2B +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 16, face = "bold"))

Fig_2A
Fig_2B

ggsave(Fig2,filename = "./Figs/Fig2_efficiency_malarone.png",units = "px", width = 6000, height = 3000)

@@

## Testing for Treatment effect on parasitemia change between first and second sampling
## Note: 
## intensity change -100 codes for individuals that were uninfected throughout 
## the experiment!
## -----------------------------------------------------------------------------
lmerTest::lmer(intensity_change ~ Treatment + Year + days_resample + (1|Nest), 
               data = df[!df$intensity_change %in% -100,]) %>%
  summary()
```

@@

```{r}
## Table S1
## -----------------------------------------------------------------------------
lm_d1 = lm(Breathing_rate ~ Treatment, data = df)
summary(lm_d1)
```

@@

```{r}
## Table S1
## -----------------------------------------------------------------------------
lm_d2 <- lm(body_temperature ~ Treatment, data = df)
summary(lm_d2)
```

@@

```{r}
## Table S1
## -----------------------------------------------------------------------------
lm_d3 <- lm(delta_body_condition~Treatment, data = df)
summary(lm_d3)
```

@@

```{r}
## Table S1
## -----------------------------------------------------------------------------
lm_d4 <- lm(Growth_rate~Treatment, data = df)
summary(lm_d4)

@@

Exploring differences in physiological measures that are associated with the infection stage (increase, peak or decreasing) of tested individuals. All models account for a number of fixed effects and non-independence among nest mates.

@@

## Plotting predicted values across infection stages
## -----------------------------------------------------------------------------
Fig_4A <- ggplot(mod_br) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_br$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Breathing rate") + 
  ylim(c(min(mod_br$conf.low) - 1.5,max(mod_br$conf.high) + 1.5)) + 
  geom_errorbar(aes(ymin = mod_br$conf.low,ymax = mod_br$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br$predicted - mod_br$std.error, ymax = mod_br$predicted + mod_br$std.error),
                size = 8, colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        plot.margin = margin(0, 0, 0, 0, "cm"), 
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
  aes(x = 1, y = max(mod_br$conf.high) + 1, xend = 4, yend = max(mod_br$conf.high) + 1)) + 
  geom_text(aes(x = 2.5, y = max(mod_br$conf.high) + 1.5),
  label = "n.s.", size = 5)
```

@@

# Plotting predicted values across infection stages
Fig_4B <- ggplot(mod_bt) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_bt$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Body temperature") + 
  geom_errorbar(aes(ymin = mod_bt$conf.low,ymax = mod_bt$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt$predicted - mod_bt$std.error,ymax = mod_bt$predicted + mod_bt$std.error),
                size = 8,colour = wes_palette("Zissou1",4), width = 0) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title.y = element_text(size = 14, margin = unit(c(0,3,0,0),"mm")), 
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 40.6, xend = 4, yend = 40.6)) + 
    geom_text(aes(x = 2.5, y = 40.7),
              label = "n.s.", size = 5)
```

@@

Fig_4C = ggplot(mod_bc) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_bc$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Δ Body condition") + 
  geom_errorbar(aes(ymin = mod_bc$conf.low,ymax = mod_bc$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bc$predicted - mod_bc$std.error,ymax = mod_bc$predicted + mod_bc$std.error),
                size = 8,colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bc$conf.high) + 17, xend = 4, yend = max(mod_bc$conf.high) + 17)) + 
  geom_text(aes(x = 2.5, y = max(mod_bc$conf.high) + 23),
              label = "n.s.", size = 5)
```

@@

```{r mixed model growth rate}
# Model 
lmer_gr = lmer(Growth_rate ~ Diff_Inf_Class + Diff_Inf_Class:Treatment + Sex +
                 days_resample + Age_in_days + Year + (1|Nest), data = df)

@@

Fig_4D = ggplot(mod_gr) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_gr$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Growth rate") + 
  geom_errorbar(aes(ymin = mod_gr$conf.low,ymax = mod_gr$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_gr$predicted - mod_gr$std.error,ymax = mod_gr$predicted + mod_gr$std.error), 
                size = 8,colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none", 
        axis.title.y = element_text(size = 14, margin = unit(c(0, 3, 0, 0), "mm")),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = margin(0, 0, 0, 30, "pt"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_gr$conf.high) + 1.2, xend = 4, yend = max(mod_gr$conf.high) + 1.2)) + 
    geom_text(aes(x = 2.5, y = max(mod_gr$conf.high) + 2),
              label = "n.s.", size = 5)
```

@@

The code below is used to summarise the model output and creates Figure 4 and Table 3 as shown in the manuscript.

```{r}
## Figure 4 
## -----------------------------------------------------------------------------
fig4 <- Fig_4A + Fig_4B + Fig_4C + Fig_4D + plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 16, face = "bold"))
fig4

ggsave(fig4,filename = "./Figs/Fig4_predicted_physio_param_infection_stages.png",units = "px", width = 3700, height = 3700)

@@

Exploration of effects by parasitemia levels on physiology of individuals. All models account for a number of fixed effects and non-independence among nest mates.

@@

```{r mixed model breathing rate 2}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_br_para = lmer(
  Breathing_rate ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample + 
    Age_in_days + Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_br_para_no_intercept = lmer(
  Breathing_rate ~ 0 + Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample + 
    Age_in_days + Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest), data = df)

## Obtain effect sizes for post-treatment parasitemia and derive model predictions
## -----------------------------------------------------------------------------
mod_br_para <- 
  left_join(plot_model(lmer_br_para_no_intercept, se = F, )[["data"]][1:5,], ## 95% CI
            plot_model(lmer_br_para_no_intercept, se = T, )[["data"]][1:5,]) %>%  ## SE
  mutate(predicted = estimate) %>% 
  mutate(lower_CI = conf.low) %>% 
  mutate(higher_CI = conf.high) %>% 
  mutate(se = std.error) %>% 
  mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

## Obtain sample sizes for labels 
## ----------------------------------------------------------------------------- 
m1 = table(df$Infection_intensity.2[!is.na(df$Breathing_rate)])
mod_br_para <- mod_br_para %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",m1)), levels = c(paste0(x,"\nn= ",m1))))

# Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_5A = ggplot(mod_br_para) + 
  aes(as.factor(mod_br_para$x),mod_br_para$predicted) + 
  geom_hline(yintercept = mean(mod_br_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Breathing rate") + 
  geom_errorbar(aes(ymin = mod_br_para$lower_CI,
                    ymax = mod_br_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br_para$predicted - mod_br_para$se, 
                    ymax = mod_br_para$predicted + mod_br_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_br_para$higher_CI, na.rm = T) + 1.2,
                              xend = 5, yend = max(mod_br_para$higher_CI, na.rm = T) + 1.2)) + 
    geom_text(aes(x = 3, y = max(mod_br_para$higher_CI, na.rm = T) + 2),
              label = "n.s.", size = 5)
```

```{r  mixed model body temperature 2, warning=FALSE, message=F, echo=T}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_bt_para = lmer(
  body_temperature ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Ambient_Temperature_AvgC + Age_in_days + Weight_g + Sex + Year + (1|Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_bt_para_no_intercept = lmer(
  body_temperature ~ 0 + Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Ambient_Temperature_AvgC + Age_in_days + Weight_g + Sex + Year + (1|Nest), data = df)

## Obtain effect sizes for post-treatment parasitemia and derive model predictions
## -----------------------------------------------------------------------------
mod_bt_para <- 
  left_join(plot_model(lmer_bt_para_no_intercept, se = F, )[["data"]][1:5,], ## 95% CI
            plot_model(lmer_bt_para_no_intercept, se = T, )[["data"]][1:5,]) %>%  ## SE
  mutate(predicted = estimate) %>% 
  mutate(lower_CI = conf.low) %>% 
  mutate(higher_CI = conf.high) %>% 
  mutate(se = std.error) %>% 
  mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

## Obtain sample sizes for labels 
## ----------------------------------------------------------------------------- 
m2 = table(df$Infection_intensity.2[!is.na(df$body_temperature)])

## Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_5B = ggplot(mod_bt_para) + 
  aes(as.factor(mod_bt_para$x),mod_bt_para$predicted) + 
  geom_hline(yintercept = mean(mod_bt_para$predicted, na.rm = T), linetype = "dashed") + 
  geom_errorbar(aes(ymin = mod_bt_para$lower_CI,ymax = mod_bt_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt_para$predicted - mod_bt_para$se,ymax = mod_bt_para$predicted + mod_bt_para$se),
                size = 8,colour = wes_palette("Zissou1",5), width = 0,) + 
  labs(x = "", y = "Body temperature") + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                              xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2)) + 
  geom_text(aes(x = 3, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.25),
            label = "0.006", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                   xend = 1, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.05)) + 
  geom_segment(size = .8,
               aes(x = 5, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.05,
                   xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2))
```

```{r mixed model body condition change 2}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_bc_para <- lmer(
  delta_body_condition ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Age_in_days + Sex + Year + (1|Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_bc_para_no_intercept <- lmer(
  delta_body_condition ~ 0+ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Age_in_days + Sex + Year + (1|Nest), data = df)

## Obtain effect sizes for post-treatment parasitemia and derive model predictions
## -----------------------------------------------------------------------------
mod_bc_para <- 
  left_join(plot_model(lmer_bc_para_no_intercept, se = F, )[["data"]][1:5,], ## 95% CI
            plot_model(lmer_bc_para_no_intercept, se = T, )[["data"]][1:5,]) %>%  ## SE
  mutate(predicted = estimate) %>% 
  mutate(lower_CI = conf.low) %>% 
  mutate(higher_CI = conf.high) %>% 
  mutate(se = std.error) %>% 
  mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

# Table with sample sizes specific for body condition change data
m3 = table(df$Infection_intensity.2[!is.na(df$delta_body_condition)])
mod_bc_para <- mod_bc_para %>% 
  mutate(x = factor(c(paste0(x,"\nn= ",m3)), levels = c(paste0(x,"\nn= ",m3))))

@@

```{r mixed model growth rate 2}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_gr_para = lmer(
  Growth_rate ~ Infection_intensity.2 + Infection_intensity.2:Treatment +
    days_resample + Sex + Age_in_days + Year + (1 | Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_gr_para_no_intercept  = lmer(
  Growth_rate ~ 0 + Infection_intensity.2 + Infection_intensity.2:Treatment +
    days_resample + Sex + Age_in_days + Year + (1 | Nest), data = df)

## Obtain effect sizes for post-treatment parasitemia and derive model predictions
## -----------------------------------------------------------------------------
mod_gr_para <- 
  left_join(plot_model(lmer_gr_para_no_intercept, se = F, )[["data"]][1:5,], ## 95% CI
            plot_model(lmer_gr_para_no_intercept, se = T, )[["data"]][1:5,]) %>%  ## SE
  mutate(predicted = estimate) %>% 
  mutate(lower_CI = conf.low) %>% 
  mutate(higher_CI = conf.high) %>% 
  mutate(se = std.error) %>% 
  mutate(x = factor(c("0","1","2","3","4"), levels = c("0","1","2","3","4")))

# Table with sample sizes specific for growth rate data
m4 = table(df$Infection_intensity.2[!is.na(df$Growth_rate)])
